Governing through patronage: the bargain for education in decentralized Brazil

### 2020-09-10 paper/paper.tex
continue to redraft paper

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\renewcommand{\theproposition}{\arabic{proposition}}
\newcommand{\real}{\mathbb{R}_+^n}
\newcommand{\bfv}{\mathbf{v}}
\newcommand{\blue}{\textcolor{blue}}

@@

\subsection{Education in Brazil}

@@

\begin{figure}[h]
    \includegraphics{plots/descriptive_enrollment.pdf}
    \caption{enrollment}
    \label{fig:enroll}
    \centering
\end{figure}

The local executive branch has sole jurisdiction over the hiring and firing of teachers and school principals. School principals are considered positions of trust (\textit{cargos de confiança}) and usually appointed by the mayor or secretary of education \citet{brollo_victor_2017}. Teachers usually enter the public service through a public exam and are eligible for tenure after two years \citet{gatti_formacao_2010}. However municipalities increasingly resort to temporary contracts to hire new teachers as budgetary pressures amount. Municipal teachers and school principals are overseen by a local department of education \citet{militao_educacao_2014}.

@@

\begin{figure}[h]
    \includegraphics{plots/saeb_map.pdf}
    \caption{hello}
    \label{fig:saeb_map}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics{plots/descriptive_staff_turnover.pdf}
    \label{fig:staff_turnover}
    \caption{hello}
\end{figure}

@@

\begin{figure}
    \includegraphics{plots/hist_coalition.pdf}
    \caption{hello}
    \label{fig:hist_coalition}
\end{figure}

>>>>

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\renewcommand{\theproposition}{\arabic{proposition}}
\newcommand{\real}{\mathbb{R}_+^n}
\newcommand{\bfv}{\mathbf{v}}
\newcommand{\blue}{\textcolor{blue}}
% \renewcommand{\includegraphics}{\includegraphics[width=0.6\linewidth]}

@@

\subsection{Municipal education in Brazil}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/descriptive_enrollment.pdf}
    \caption{enrollment}
    \label{fig:enroll}
\end{figure}

The local executive branch has sole jurisdiction over the hiring and firing of teachers and school principals. School principals are considered positions of trust (\textit{cargos de confiança}) and usually appointed by the mayor or secretary of education \citet{brollo_victor_2017}. Teachers usually enter the public service through a public exam and are eligible for tenure after two years \citet{gatti_formacao_2010}. However municipalities increasingly resort to temporary contracts to hire new teachers as budgetary pressures amount. Municipal teachers and school principals are overseen by a local department of education \citep{militao_educacao_2014}.

Positions in the educational sector are politically valuable. As a department, it comprises over 25 percent of local public sector jobs. Due to their relatively high compensation and social status compared to other positions in the local bureaucracy, educational staff positions are particularly valuable for constituencies seeking employment. Additionally, anecdotal evidence collected during fieldwork suggests that teachers play an important role in local party networks, leading electoral campaigns and brokering votes, similar to dynamics found by \citet{oliveros_making_2016} in Argentina's campaigning teachers.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/descriptive_staff_breakdown}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/saeb_map.pdf}
    \caption{Uneven quality of municipal education. Polygons represent municipalities, warmer colors indicate higher municipal average in standardized test scores. Averages are weighted by student participation rate.}
    \label{fig:saeb_map}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.\linewidth]{plots/descriptive_staff_turnover.pdf}
    \label{fig:staff_turnover}
    \caption{Proportion of public sector jobs by department. Note that I only keep the top 5 categorie}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{plots/hist_coalition.pdf}
    \caption{Distribution of share of legislative seats allied with the mayor during the electoral campaign for the elecoral period of 2002-2016.}
    \label{fig:hist_coalition}
\end{figure}
